Concretizing the Process Hitting into Biological Regulatory Networks

### 2013-04-17 parts/grn.tex
add unsigned edges to parameters inference
more explicit \new labels

\begin{definition}[\new{Interaction Graph}]
\label{def:ig}
%An \emph{Interaction Graph} (IG) is a quadruplet $(\Gamma, E_+, E_-, E_\uns)$ where $\Gamma$ is a finite number of \emph{components},
%and $E_+$ (resp. $E_-$) $\subset \{a \xrightarrow{t} b \mid a, b \in \Gamma \wedge t \in [1; l_a]\}$
%is the set of positive (resp. negative) \emph{regulations} between two nodes, labeled with a \emph{threshold}.
An \emph{Interaction Graph} (IG) is a quadruplet $(\Gamma, E)$ where $\Gamma$ is a finite number of \emph{components},
and $E \subset \{ \GRNedge{a}{s}{t}{b} %\in \Gamma \times S \times \sN \times \Gamma 
\mid a, b \in \Gamma \wedge s \in \{ +, -, \uns \} \wedge t \in [1; l_a]\}$
%and $E_+$ (resp. $E_-$) $\subset \{a \xrightarrow{t} b \mid a, b \in \Gamma \wedge t \in [1; l_a]\}$
is the set of 
%positive (resp. negative)
\emph{regulations} between two nodes, labeled with a \emph{sign} and a \emph{threshold}.

@@

\begin{definition}[\new{Effective levels ($\levelsl$)}]\label{def:levels}
Let $(\Gamma,E)$ be an IG and $a, b \in \Gamma$ two of its components:
\begin{itemize}
%  \item if $a \xrightarrow{t} b \in E_+$, $\levelsA{a}{b} \DEF [t; l_a]$ and
%    $\levelsI{a}{b} \DEF [0; t-1]$;
%  \item if $a \xrightarrow{t} b \in E_-$, $\levelsA{a}{b} \DEF [0; t-1]$ and
%    $\levelsI{a}{b} \DEF [t; l_a]$;
%  \item otherwise, $\levelsA{a}{b} \DEF \levelsI{a}{b} \DEF \emptyset$.
  \item if $\GRNedge{a}{s}{t}{b} \in E$, $\levels{a}{b} \DEF [t; l_a]$;
  \item otherwise, $\levels{a}{b} \DEF \emptyset$.
\end{itemize}
\end{definition}

@@

@@

\begin{definition}[\new{Asynchronous dynamics}]\label{def:dynamics}
The dynamics of a BRN using Thomas' parameters is defined by the transition relation $\GRN{}{} \in \prod_{a \in \Gamma} [0;l_a] \times \prod_{a \in \Gamma} [0;l_a]$:
$\forall s, s' \in \prod_{a \in \Gamma} [0;l_a]$,
\begin{align*}
\GRNtrans{s}{s'} &\Leftrightarrow \exists a \in \Gamma, \GRNget{s}{a} \notin K_{a, \GRNres{a}{s}} \wedge \GRNget{s'}{a} = \GRNget{s}{a} + f^a(s) \\
  & \qquad \wedge \forall b \in \Gamma, b \neq a \Rightarrow \GRNget{s}{b} = \GRNget{s'}{b}
\end{align*}
with: $f^a(s) = 
  \begin{cases}
    +1 & \text{if } \GRNget{s}{a} < K_{a, \GRNres{a}{s}} \\
    -1 & \text{if } \GRNget{s}{a} > K_{a, \GRNres{a}{s}} \\
  \end{cases}$
\begin{comment}
Let $s$ be a state of a BRN using Thomas' parameters $(\IG, K)$ where $\IG = (\Gamma, E_+, E_-)$.
The state that succeeds to $s$ is given by the indeterministic function $f(s)$:
\begin{align*}
  f(s)  & = s' \Leftrightarrow \exists a \in \Gamma,
    \GRNget{s'}{a} = f^a(s) \wedge
    \forall b \in \Gamma, b \neq a, \GRNget{s}{b} = \GRNget{s'}{b}
    \quad\text{, with}\\
  f^a(s) & =
  \begin{cases}
    \GRNget{s}{a} + 1 & \text{if } \GRNget{s}{a} < K_{a,A,B} \\% \GRNres{a}{s}} \\
    \GRNget{s}{a} & \text{if } \GRNget{s}{a} \in K_{a,A,B} \\ %\GRNres{a}{s}}\\
    \GRNget{s}{a} - 1 & \text{if } \GRNget{s}{a} > K_{a,A,B} % \GRNres{a}{s}}
  \end{cases}
\quad\text{, where $A,B=\GRNres{a}{s}$.}
\end{align*}
\end{comment}
\end{definition}

>>>>

\new{Added unsigned edges ($\uns$) that stand for an unknown king of regulation.
The definition of $\levels{a}{b}$ now gives the levels for which $a$ regulates $b$ (i.e. $a > t$), independently from the regulation sign.}

\begin{definition}[Interaction Graph]
\label{def:ig}
%An \emph{Interaction Graph} (IG) is a quadruplet $(\Gamma, E_+, E_-, E_\uns)$ where $\Gamma$ is a finite number of \emph{components},
%and $E_+$ (resp. $E_-$) $\subset \{a \xrightarrow{t} b \mid a, b \in \Gamma \wedge t \in [1; l_a]\}$
%is the set of positive (resp. negative) \emph{regulations} between two nodes, labeled with a \emph{threshold}.
An \emph{Interaction Graph} (IG) is a quadruplet $(\Gamma, E)$ where $\Gamma$ is a finite number of \emph{components},
and $E \subset \{ \GRNedge{a}{s}{t}{b} %\in \Gamma \times S \times \sN \times \Gamma 
\mid a, b \in \Gamma \wedge s \in \{ +, -, \uns \} \wedge t \in [1; l_a]\}$
%and $E_+$ (resp. $E_-$) $\subset \{a \xrightarrow{t} b \mid a, b \in \Gamma \wedge t \in [1; l_a]\}$
is the set of 
%positive (resp. negative)
\emph{regulations} between two nodes, labeled with a \emph{sign} and a \emph{threshold}.

@@

\begin{definition}[Effective levels ($\levelsl$)]\label{def:levels}
Let $(\Gamma,E)$ be an IG and $a, b \in \Gamma$ two of its components.
If $\GRNedge{a}{s}{t}{b} \in E$, we define: $\levels{a}{b} \DEF [t; l_a]$ and $\ulevels{a}{b} \DEF [0; t-1]$.
%%%
%\begin{itemize}
%  \item if $a \xrightarrow{t} b \in E_+$, $\levelsA{a}{b} \DEF [t; l_a]$ and
%    $\levelsI{a}{b} \DEF [0; t-1]$;
%  \item if $a \xrightarrow{t} b \in E_-$, $\levelsA{a}{b} \DEF [0; t-1]$ and
%    $\levelsI{a}{b} \DEF [t; l_a]$;
%  \item otherwise, $\levelsA{a}{b} \DEF \levelsI{a}{b} \DEF \emptyset$.
%%%
%  \item if $\GRNedge{a}{s}{t}{b} \in E$, $\levels{a}{b} \DEF [t; l_a]$;
%  \item otherwise, $\levels{a}{b} \DEF \emptyset$.
%\end{itemize}
\end{definition}

@@

\new{No more $A,B$ but sets $\omega$ of resources instead. Definitions of $K$ were adapted.}

@@

\begin{definition}[Asynchronous dynamics]\label{def:dynamics}
The dynamics of a BRN using Thomas' parameters is defined by the transition relation $\GRNtrans{}{} \in \prod_{a \in \Gamma} [0;l_a] \times \prod_{a \in \Gamma} [0;l_a]$:
$\forall s, s' \in \prod_{a \in \Gamma} [0;l_a]$,
\begin{align*}
\GRNtrans{s}{s'} &\Leftrightarrow \exists a \in \Gamma, \GRNget{s}{a} \notin K_{a, \GRNres{a}{s}} \wedge \GRNget{s'}{a} = \GRNget{s}{a} + f^a(s) \\
  & \qquad \wedge \forall b \in \Gamma, b \neq a \Rightarrow \GRNget{s}{b} = \GRNget{s'}{b}
\end{align*}
with: $f^a(s) = 
  \begin{cases}
    +1 & \text{if } \GRNget{s}{a} < K_{a, \GRNres{a}{s}} \\
    -1 & \text{if } \GRNget{s}{a} > K_{a, \GRNres{a}{s}} \\
  \end{cases}$
\begin{comment}
Let $s$ be a state of a BRN using Thomas' parameters $(\IG, K)$ where $\IG = (\Gamma, E_+, E_-)$.
The state that succeeds to $s$ is given by the indeterministic function $f(s)$:
\begin{align*}
  f(s)  & = s' \Leftrightarrow \exists a \in \Gamma,
    \GRNget{s'}{a} = f^a(s) \wedge
    \forall b \in \Gamma, b \neq a, \GRNget{s}{b} = \GRNget{s'}{b}
    \quad\text{, with}\\
  f^a(s) & =
  \begin{cases}
    \GRNget{s}{a} + 1 & \text{if } \GRNget{s}{a} < K_{a,A,B} \\% \GRNres{a}{s}} \\
    \GRNget{s}{a} & \text{if } \GRNget{s}{a} \in K_{a,A,B} \\ %\GRNres{a}{s}}\\
    \GRNget{s}{a} - 1 & \text{if } \GRNget{s}{a} > K_{a,A,B} % \GRNres{a}{s}}
  \end{cases}
\quad\text{, where $A,B=\GRNres{a}{s}$.}
\end{align*}
\end{comment}
\end{definition}

### 2013-04-17 parts/infer-param.tex
more explicit \new labels

This subsection presents some results related to the inference of independent discrete parameters from a given PH. These results are equivalent to those presented in \cite{PMR10-TCSB}, with notation adapted to be shared with the previous section.
In addition, we introduce the well-formed PH for parameter inference property (\pref{pro:wf-ph-K}),
which implies that the inferred IG does not contain any unsigned interactions, and thus can be seen as the
regular IG $(\Gamma, E_+, E_-)$,
and that any processes in $\levelsA{b}{a}$ (resp. $\levelsI{b}{a}$) share the same behavior
regarding $a$.
\begin{property}[\new{Well-formed PH for parameter inference}]\label{pro:wf-ph-K}
A PH is well-formed for parameter inference if and only if
it is well-formed for IG inference, and
the IG $(\Gamma, E_+, E_-, E_\pm)$ inferred by \pref{pps:inference-IG}
verifies $E_\pm=\emptyset$ and if the following property holds:
\begin{align}
\begin{split}
\forall b\in \GRNreg{a} &,
	\forall (i,j\in\levels{b}{a} \vee i,j\in\ulevels{b}{a}), \\
& \qquad	\forall c, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a} \wedge b\in\PHdirectpredec{c})), \\
& \qquad \qquad
			\PHfrappe{b_i}{c_k}{c_l}\in\PHa \Leftrightarrow
				\PHfrappe{b_j}{c_k}{c_l}\in\PHa
\end{split}
\label{eq:wf-levels}
\end{align}
\end{property}

@@

>>>>

This subsection presents some results related to the inference of independent discrete parameters from a given PH. These results are equivalent to those presented in \cite{PMR10-TCSB}, with notation adapted to be shared with the previous section.
In addition, we introduce the well-formed PH for parameter inference property (\pref{pro:wf-ph-K}),
which implies that the inferred IG does not contain any unsigned interactions, and thus can be seen as the
regular IG $(\Gamma, E_+, E_-)$,
and that any processes in $\levelsA{b}{a}$ (resp. $\levelsI{b}{a}$) share the same behavior
regarding $a$.

\new{Changed $\levelsA{a}{b}$ and $\levelsI{a}{b}$ into $\levels{a}{b}$ in \pref{eq:wf-levels} and \pref{eq:param_context}.}

\begin{property}[Well-formed PH for parameter inference]\label{pro:wf-ph-K}
A PH is well-formed for parameter inference if and only if
it is well-formed for IG inference, and
the IG $(\Gamma, E_+, E_-, E_\pm)$ inferred by \pref{pps:inference-IG}
verifies $E_\pm=\emptyset$ and if the following property holds:
\begin{align}
\begin{split}
\forall b\in \GRNreg{a} &,
	\forall (i,j\in\levels{b}{a} \vee i,j\in\ulevels{b}{a}), \\
& \qquad	\forall c, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a} \wedge b\in\PHdirectpredec{c})), \\
& \qquad \qquad
			\PHfrappe{b_i}{c_k}{c_l}\in\PHa \Leftrightarrow
				\PHfrappe{b_j}{c_k}{c_l}\in\PHa
\end{split}
\label{eq:wf-levels}
\end{align}
\end{property}

@@

\new{Update all assumptions.}
